---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# trapgridR

<!-- badges: start -->
<!-- badges: end -->

trapgridR is an interace for the java program "TrapGrid" (Manoukis et al. 2015) using R. Calling this program from R allows for multiple simulations to be run, and the results to be easily collated and summarised.




## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("mattecologist/trapgridR")
```
## Example

```{r setup}
library(trapgridR)
```

There is a simple function to make a regular grid across a given area
```{r figure.width=4, figure.height=4}
make_regular_grid("my_grid")
```

Simulations can be setup in R and results stored as objects
```{r}
model1 <- trapgridR(filepath="my_grid", nDays = 14, nFlies = 100, nSim=10, D=10^5)
```

Simulation results can be plotted from the model objects
```{r plot of simulation results, figure.width=4, figure.height=3}
library (ggplot2)
ggplot(model1$simRuns, aes(Day, 1-Cumulative.Escape.Probability, colour=SimRun, group=SimRun))+
  geom_line()+
  theme_dark()+
  scale_x_continuous(expand=c(0.01,0))+
  scale_y_continuous(expand=c(0,0.01))

```

Additionally, fly locations from the simulation can be plotted to examine movement patterns....
```{r figure.width=4, figure.height=4}
ggplot (model1$flyLoc, aes(as.integer(X), as.integer(Y),  colour=as.integer(Simulation.Number)))+
  geom_point()+
  facet_wrap(.~as.integer(Day))+
  scale_colour_viridis_c()+
  theme_dark()
```




.....which is a bit messy, so using `{gganimate}` can make an animation:

```{r figure.width=3, figure.height=3}
library (gganimate)
ggplot (model1$flyLoc, aes(as.integer(X), as.integer(Y),  colour=as.integer(Simulation.Number)))+
  geom_point()+
  transition_time(as.integer(Day)) +
  scale_colour_viridis_c()+
  theme_dark()

```

We've included functions so that actualy trap arrangements can be easily be used in R, and are implementing further changes to the model that allow specification of further biological and trapping parameters.

# References

Manoukis, Nicholas C., Brian Hall, and Scott M. Geib. “A Computer Model of Insect Traps in a Landscape.” Scientific Reports 4, no. 1 (May 2015). https://doi.org/10.1038/srep07015.
