---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# trapgridR

<!-- badges: start -->
<!-- badges: end -->

trapgridR is an interace for the java program "TrapGrid" (Manoukis et al. 2015) using R. Calling this program from R allows for multiple simulations to be run, and the results to be easily collated and summarised.


## Installation

Prior to installation, make sure you have `{rJava}` installed and working. This may require installation of Java first. For example on Ubuntu:

`sudo apt-get install -y default-jre`

`sudo apt-get install -y default-jdk`

`sudo R CMD javareconf`


You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("mattecologist/trapgridR")
```



## Example

```{r setup}
library (trapgridR)
library (ggplot2)
library (sp)
library (raster)
library (maptools)
```

At the most basic level, there is a simple function to make a regular grid across a given area
```{r fig.width=5, fig.height=5}
make_regular_grid("my_grid", #name of file to write out for TrapGrid
  gridSize = c(364, 572), # upper right corner coordinates (lower left is always 0, 0)
  gridSpace = 80, # spacing of traps in metres
  lambda = 0.02) # lambda of the traps
```
Or make a random grid of at least a minimum distance apart, that can include additional perimeter traps or not

```{r fig.width=5, fig.height=5}
make_random_grid(n.traps=5, gridname="my_grid", perim=TRUE)
```

## Simulating Orchards / Management areas

Alternatively, you might have a polygon/shapefile of an orchard you want to explore different scenarios over. You can also build randomly shaped and sized polygons to investigate. We were interseted in orchard "blocks" so have used that term to describe them, but it can be any management area that is appropriate.

```{r}
set.seed(3056)

new.block <- make_random_block(n.sides = 4,         # specify the number of sides to the block 
                               regular.block = F,   # specify if all sides equal or random
                               block.size = 1e+05)  # specify the size of the block in metres squared

new.traps <- make_random_traps(block=new.block, n.traps = 5, min.dist = 50) # add traps to the block - here just 4 random traps all at least 50 meters apart


plot (new.block) # the traps object stores the new block polygon too
points (new.traps[[1]], pch=20) # the traps are here as point

```

# Setting outbreak locations

Its possible to take the block polygon and seed all the outbreaks either within or outside of the block
```{r}
set.seed (5082)

outbreaks <- make_block_outbreak(traps= new.traps[1], 
                                 block = new.traps[[2]],
                                 nOutbreaks = 10,    # how many outbreak locations 
                                 outbreak_buf = 500) # buffer around the orchard for outbreaks to seeded in

plot (outbreaks$buffer_region)
points (outbreaks$outbreak_set)
points (outbreaks$traps_sp, pch=20)
```
```{r}
make_block_grid(gridname = "orchard_traps",
                traps = outbreaks$traps_sp,
                lambda = 0.02) # trap attractiveness value (1/0.02 = 50 metres for 65% capture probability)
```

```{r}
model.temp <- trapgridR("orchard_traps", nDays=14, nFlies=50, D=10^4,
                        outbreaks = "outbreaks")

```


Simulation results can be plotted from the model objects. 
```{r plot of simulation results, fig.width=5, fig.height=5}
library (ggplot2)
ggplot(model.temp$simRuns, aes(Day, 1-Cumulative.Escape.Probability))+
  geom_line(aes(group=SimRun), colour=viridis::viridis(3)[1], alpha=0.3)+
  geom_smooth(method="gam", colour=viridis::viridis(3)[2])+
  theme_dark()+
  scale_x_continuous(expand=c(0.01,0))+
  scale_y_continuous(expand=c(0,0.01))+
  theme_minimal()

```
The 1-cumulative escape probability can be thought of as the detection probability.

Additionally, fly locations from the simulation can be plotted to examine the diffusion process across the landscape over time.
```{r fig.width=8, fig.height=4}
ggplot (model.temp$flyLoc, aes(as.integer(X), as.integer(Y),  colour=as.factor(Simulation.Number)))+
  geom_point()+
  facet_wrap(.~as.integer(Day))+
  scale_colour_viridis_d()+
  theme_minimal()
```


We've included functions so that actualy trap arrangements can be easily be used in R, and are implementing further changes to the model that allow specification of further biological and trapping parameters.

# References

Manoukis, Nicholas C., Brian Hall, and Scott M. Geib. “A Computer Model of Insect Traps in a Landscape.” Scientific Reports 4, no. 1 (May 2015). https://doi.org/10.1038/srep07015.
